---
title: "Untitled"
format: docx
editor: visual
---
#Loading all the necessary packages
```{r, echo=FALSE}
library(readr)
library(sqldf)
library(mosaic)
library(gtsummary)
library(gt)
library(corrplot)
library(table1)
library(tigerstats)
library(ggplot2)
library(tidyverse)
library(psych)
```
#Loading all the datasets
```{r, echo=FALSE}
BenchmarkData <- read_csv("R-projects/Annual Employee Survey Analysis/BenchmarkData.csv",show_col_types = FALSE)
TrainingData <- read_csv("R-projects/Annual Employee Survey Analysis/TrainingData.csv",show_col_types = FALSE)
PersonalData <- read_csv("R-projects/Annual Employee Survey Analysis/PersonalData.csv",show_col_types = FALSE)
PerformanceData <- read_csv("R-projects/Annual Employee Survey Analysis/PerformanceData.csv",show_col_types = FALSE)
CompensationData <- read_csv("R-projects/Annual Employee Survey Analysis/CompensationData.csv",show_col_types = FALSE)
SelectionValidationData <- read_csv("R-projects/Annual Employee Survey Analysis/SelectionValidationData.csv",show_col_types = FALSE)
SurveyData <- read_csv("R-projects/Annual Employee Survey Analysis/SurveyData.csv",show_col_types = FALSE)
```

● Report a list of the employee IDs for those employees who responded with Strongly
Agree (5) to the following item: “During the next 12 months, I will probably look for a
new job outside of Recreation Unlimited.” (TI2).
```{r,fig.align="center", echo = FALSE,fig.width = 10}
employee_IDs <- sqldf("SELECT PersonalData.EmployeeID,SurveyData.TI2 FROM PersonalData 
                      LEFT JOIN SurveyData ON PersonalData.EmployeeID = SurveyData.EmployeeID
                      WHERE TI2 = 5")
employee_IDs|>gt()|>tab_options(column_labels.background.color = "darkorange")
```

● Report information about the internal consistency reliabilities (Cronbach’s alphas) of
each of the multi-item survey measures; explain and justify which items you selected to
create a composite variable (i.e., overall scale score variable) for each of the multi-item
measures. The executives would like you to report which items you removed from each
measure (if any) prior to creating a composite variable (i.e., overall scale score
variable). Moreover, they would like for you to justify those decisions by referencing
Cronbach’s alphas and the actual item content (see survey items above or in link to
online survey).

#*Internal consistency reliabilities (Cronbach’s alphas) of job satisfaction items*
```{r, echo=FALSE}
alpha(SurveyData[,c("JS1","JS2","JS3","JS4")])
```
The raw alpha (raw_alpha) based on job satisfaction (JS1,JS3,JS4) items falls under the questionable category in descriptors for Cronbach’s alpha while job satisfaction (JS2) exceeds the acceptable threshold of 0.70. 
- Therefore, We will retain all four items when computing the composite variable for job satisfaction because:
. Cronbach’s alpha for all four items is above 0.60 ;
. Removing an item would decrease Cronbach’s alpha; and
. The content of all 3 items seems to fit within the conceptual definition of job satisfaction.
#*Internal consistency reliabilities (Cronbach’s alphas) of engagement items*
```{r, echo=FALSE}
alpha(SurveyData[,c("Eng1","Eng2","Eng3")])
```
The raw alpha (raw_alpha) based on all three engagement items exceeds the acceptable threshold of 0.70. 
- Therefore, We will retain all three items when computing the composite variable for engagement because:
. Cronbach’s alpha for all 3 items is above 0.70 (and thus acceptable);
. Removing an item would decrease Cronbach’s alpha; and
. The content of all 3 items seems to fit within the conceptual definition of engagement.
#*Internal consistency reliabilities (Cronbach’s alphas) of turnover intentions items*
```{r, echo=FALSE}
alpha(SurveyData[,c("TI1","TI2","TI3")])
```
The raw alpha (raw_alpha) based on turnover intentions (TI1,TI2) items falls under Unacceptable category in descriptors for Cronbach’s alpha while turnover intentions (TI3) exceeds the acceptable threshold of 0.70. 
- Therefore, We will only retain TI3 item when computing the composite variable for turnover intentions because:
. its Cronbach’s alpha is above the acceptable threshold of 0.70 ;
. The content of TI3 item seems to fit within the conceptual definition of turnover intentions.

● Create composite variables (i.e., overall scale score variables) for the three multi-item
survey measures (see above request/bullet for information regarding how to
explain/justify which items to retain).
#*Internal consistency reliabilities (Cronbach’s alphas) of turnover intentions items*
```{r, echo=FALSE}
SurveyData$JS_Overall <- rowMeans(SurveyData[,c("JS1","JS2","JS3","JS4")], na.rm=TRUE)
SurveyData$Eng_Overall <- rowMeans(SurveyData[,c("Eng1","Eng2","Eng3")], na.rm=TRUE)
SurveyData$TI_Overall <- rowMeans(SurveyData[,"TI3"], na.rm=TRUE)
```

```{r, echo=FALSE}
composite_var <- SurveyData|>select(JS_Overall,Eng_Overall,TI_Overall)
```

● Report and interpret appropriate measures of central tendency and dispersion for the
three composite variables (based on their measurement scales) you created for the job
satisfaction items, engagement items, and turnover intentions items.
*Measures of central tendency and dispersion for the three composite variables*
```{r, echo=FALSE}
JS_Overall_desc = favstats(~JS_Overall, data = composite_var)
Eng_Overall_desc = favstats(~Eng_Overall, data = composite_var)
TI_Overall_desc = favstats(~TI_Overall, data = composite_var)
Composite_Variable  = c("JS_Overall","Eng_Overall","TI_Overall")
Summary <- rbind(JS_Overall_desc,Eng_Overall_desc,TI_Overall_desc)
Summary <- cbind(Composite_Variable,Summary)
Summary|>gt()|>tab_options(column_labels.background.color = "darkgreen")
```



● Compare Recreation Unlimited’s survey responses to employees’ responses from the 23
competitor organizations. The means across the 23 competitor organizations can be
found in the data file called BenchmarkData.csv.
```{r, echo=FALSE}
BenchmarkData_t <- data.frame(cbind(names(BenchmarkData),t(BenchmarkData)))
BenchmarkData_t$X2 <- as.numeric(as.character(BenchmarkData_t$X2))
BenchmarkData_t_Mean <- BenchmarkData_t|>rename(BenchmarkResp = X1,Mean = X2)
survey_long <- SurveyData|>pivot_longer(cols = (2:11),names_to = 'survey_resp',values_to = 'Rating')|>
               select(survey_resp,Rating)

survey_long_mean <- survey_long|>group_by(survey_resp)|>summarise(Mean = round(mean(Rating),2))
comparison <- sqldf("SELECT survey_long_mean.survey_resp as `Survey question`,
                      survey_long_mean.mean as `Recreation's employees mean response`,
                      BenchmarkData_t_Mean.Mean as `Benchmark's employees mean response`
                      FROM survey_long_mean 
                      INNER JOIN BenchmarkData_t_Mean 
                      ON survey_long_mean.survey_resp = BenchmarkData_t_Mean.BenchmarkResp")
comparison|>gt()
```
```{r, fig.align="center", echo = FALSE,fig.width = 14,fig.height=10}
comparison_long <- comparison|>
  pivot_longer(cols = (2:3),
               names_to = 'Vars',values_to = 'Mean_rating')
comparison_long|>
# Stacked
ggplot(aes(fill=Vars, y=Mean_rating, x=`Survey question`)) + 
  geom_bar(position="stack", stat="identity")+ theme_bw()+
  ggtitle("Recreation Unlimited’s survey responses to employees’ responses from the 23 competitor  organizations comparisoon")+
  theme(plot.title = element_text(lineheight=3, face="bold", color="black",size=16))+
  geom_text(aes(label = Mean_rating), size = 10, hjust = 0.5, vjust = 5, position = "stack")+
  theme( axis.line = element_line(colour = "darkblue", size = 1, linetype = "solid"),
         axis.text.x = element_text(face="bold", color="#993333",size=14, angle=0),
         axis.text.y = element_text(face="bold", color="#993333",size=14, angle=0))+
  theme(legend.position="top",legend.text=element_text(size=14),legend.title=element_text(size=14))
  
```


● Report the criterion-related validities of the job satisfaction, engagement, and turnover
intentions composite variables. The behavioral anchored rating scale (BARS) and sales
revenue generated (Sales) variables will serve as your two criterion variables, and they
are saved in a file called PerformancdData.csv. Specifically, please estimate, interpret,
and report six correlations involving the following combinations of variables:
1. Job satisfaction composite variable & BARS scores (BARS)
2. Job satisfaction composite variable & sales revenue generated (Sales)
3. Engagement composite variable & BARS scores (BARS)
4. Engagement composite variable & sales revenue generated (Sales)
5. Turnover intentions composite variable & BARS scores (BARS)
6. Turnover intentions composite variable & sales revenue generated (Sales)
```{r, fig.align="center", echo = FALSE,fig.width = 12}
df <- sqldf("SELECT SurveyData.EmployeeID,SurveyData.JS_Overall,SurveyData.Eng_Overall,
                      SurveyData.TI_Overall,BARS,Sales FROM SurveyData
                      INNER JOIN PerformanceData 
                      ON SurveyData.EmployeeID = PerformanceData.EmployeeID")
#head(df)|>gt()
M = cor(df[,-1])
corrplot(M, method="number",type = "lower")
```


● Make recommendations for the organization that are informed by your findings from
the previous bullet points.


● Finally, for additional reporting guidance and requirements, be sure to refer to the
grading rubric found at the end of this document.
